Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-02-20 manuscript.Rmd
change to springer layout

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
author:
  - name: Aaron Peikert
    affil: a, b, c
    email: peikert@mpib-berlin.mpg.de
  - name: Andreas M. Brandmaier
    affil: a, b, d 
    email: brandmaier@mpib-berlin.mpg.de
affiliation:
  - num: a
    address: |
      Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
  - num: b
    address: |
     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
  - num: c
    address: |
      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
  - num: d
    address: |
      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
repro:
  scripts:
    - R/plots.R
  files:
    - abstract.md
    - data/simulation.rds
    - apa.csl
  packages:
    - here
    - tidyverse
    - ggplot2
    - ggthemes
    - patchwork
    - tikzDevice
    - rticles
    - gert
    - aaronpeikert/repro@5075336
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
    - curl
    - inkscape
output: rticles::tf_article
type: ARTICLE TEMPLATE
keywords: |
  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
---

@@

<!---many researcher do not subscribe to confirmatory agenda--->
We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they pose.
However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.

@@

First, the theory itself limits theoretical risk:
Some theories simply do not make risky predictions, and preregistration will not change that.
Consider the case of a researcher contemplating the relation between two sets of variables.
Suppose each set is separately well studied, and strong theories tell the researcher how the variables within the set relate.
However, our imaginary researcher now considers the relation between these two sets.
For lack of a better theory, they assume that some relation between any variables of the two sets exists.
This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing the theory again.
The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.

@@

The materials for this article are available on [GitHub](https://github.com/aaronpeikert/bayes-prereg) [@zenodo-release].
This version was created from git commit ``r repro::current_hash(backend = "gert")``.
The manuscript is available as preprint [@no-preprint-yet] and will be submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).

>>>>

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
titlerunning: "Preregistration increases the persuasiveness of evidence"
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
author:
  - name: Aaron Peikert
    affil: a, b, c
    email: peikert@mpib-berlin.mpg.de
  - name: Andreas M. Brandmaier
    affil: a, b, d 
    email: brandmaier@mpib-berlin.mpg.de
affiliation:
  - num: a
    address: |
      Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
  - num: b
    address: |
     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
  - num: c
    address: |
      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
  - num: d
    address: |
      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
repro:
  scripts:
    - R/plots.R
  files:
    - abstract.md
    - data/simulation.rds
    - apa.csl
  packages:
    - here
    - tidyverse
    - ggplot2
    - ggthemes
    - patchwork
    - tikzDevice
    - rticles
    - gert
    - aaronpeikert/repro@5075336
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
    - curl
    - inkscape
output: rticles::springer_article
keywords:
  - preregistration
  - confirmation
  - exploration
  - hypothesis testing
  - Bayesian
  - Open Science
biblio-style: spbasic
header-includes: |
  \usepackage{amsmath}
  \usepackage{amssymb}
---

@@

<!---many researcher do not subscribe to confirmatory agenda--->
We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they pose.
However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
It is likely that empirical sciences need more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.

@@

First, the theory itself limits theoretical risk:
Some theories simply do not make risky predictions, and preregistration will not change that.
Consider the case of a researcher contemplating the relation between two sets of variables.
Suppose each set is separately well studied, and strong theories tell the researcher how the variables within the set relate.
However, our imaginary researcher now considers the relation between these two sets.
For lack of a better theory, they assume that some relation between any variables of the two sets exists.
This is not a risky prediction to make in empirical sciences, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing the theory again.
The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.

@@

The materials for this article are available on [GitHub](https://github.com/aaronpeikert/bayes-prereg) [@zenodo-release].
This version was created from git commit ``r repro::current_hash(backend = "gert")``.
The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
